---
title: "NHL Shot based Goal Scoring Model"
author: "Sasank Vishnubhatla"
date: "February 2, 2019"
output:
  html_document:
    runtime: shiny
    theme: simplex
    df_print: paged
    dev: 'svg'
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

Last Update: `r Sys.time()`

# Libraries

Before we start, let's load a few libraries.

```{r libraries}
rm(list = ls())

set.seed(100)

options(warn = -1)

library(knitr)
library(ggplot2)
library(caret)
library(doParallel)

registerDoParallel(cores = (detectCores() - 1))
```

We register all but one core so we can have a lot of parallelsism when we start training our models.

# Data Loading

Let's read in our data.

```{r read_csv}
data.2015 = read.csv("data/2015.csv")
data.2016 = read.csv("data/2016.csv")
data.2017 = read.csv("data/2017.csv")
data.2018 = read.csv("data/2018.csv")
```

Now, we will only deal with regular season events. So let's remove the playoffs from our datasets.

```{r remove_playoffs}
get.regular.season = function(data) {
	subset(data, isPlayoffGame == 0)
}

season.2015 = get.regular.season(data.2015)
season.2016 = get.regular.season(data.2016)
season.2017 = get.regular.season(data.2017)
season.2018 = get.regular.season(data.2018)
```

Now let's remove extraneous columns. At the end, we will have the following columns (I've changed their names for ease):

| Old Column Name | New Column Name |
| --------------- | --------------- |
| `xCordAdjusted` | `x` |
| `yCordAdjusted` | `y` |
| `shotAngleAdjusted` | `angle` |
| `shotDistance` | `dist` |
| `goal` | `goal` |

```{r get_helpful}
get.helpful.data = function(data) {
	data.frame(x = data$xCordAdjusted,
		   y = data$yCordAdjusted,
		   angle = data$shotAngleAdjusted,
		   dist = data$shotDistance,
		   team = data$teamCode,
		   goal = data$goal)
}

analysis.2015 = get.helpful.data(season.2015)
analysis.2016 = get.helpful.data(season.2016)
analysis.2017 = get.helpful.data(season.2017)
analysis.2018 = get.helpful.data(season.2018)
```

Sometimes, there is incomplete data. Let's just keep all the complete cases and remove the incomplete ones.

```{r remove_incomplete}
analysis.2015 = analysis.2015[complete.cases(analysis.2015),]
analysis.2016 = analysis.2016[complete.cases(analysis.2016),]
analysis.2017 = analysis.2017[complete.cases(analysis.2017),]
analysis.all = rbind(analysis.2017, rbind(analysis.2016, analysis.2015))
analysis.all = analysis.all[complete.cases(analysis.all),]
analysis.2018 = analysis.2018[complete.cases(analysis.2018),]
```

We'll need a function to get team data.

```{r get_team_data}
get.team.data = function(data, code) {
	subset(data, team == code)
}
```

# Creating the Models

With our data, we can start creating models. We'll be creating the following models:

- Neural Network
- K Nearest Neighbors

```{r train_models}
control = trainControl(method = "repeatedcv", number = 5, repeats = 2)

model.nnet = train(goal ~ . -goal -team,
                   data = analysis.all,
                   method = "nnet",
                   trControl = control)

model.knn = train(goal ~ . -goal -team,
				  data = analysis.all,
				  method = "knn",
				  trControl = control)
```

# Extracting Predictions

Our predictions will come from `analysis.2018`. Here's what a little bit of that data looks like:

```{r testing_data}
analysis.2018
```

Now, we can use the `predict` function to get our predictions.

```{r get_predictions}
nnet.prediction = predict(model.nnet, newdata = analysis.2018)
knn.prediction = predict(model.knn, newdata = analysis.2018)

nnet.prediction.data = data.frame(analysis.2018)
nnet.prediction.data$predict = nnet.prediction

knn.prediction.data = data.frame(analysis.2018)
knn.prediction.data$predict = knn.prediction
```

So, our Neural Network data looks like:

```{r nnet_data}
nnet.prediction.data
```

Our K-Nearest Neighbors data looks like:

```{r knn_data}
knn.prediction.data
```

# Visualizing the Predictions

With our predictions, let's view how they differ. To make our lives easier, here's a few plotting functions.

```{r vis_funs}
make.knn.plot.dist = function(data, primary, secondary, team) {
	name = paste(team, "Predicted Goal Probability from KNN Model", sep = " ")
	plot = ggplot(data) +
	geom_hex(aes(x = dist, y = predict, alpha = ..count..),
			 fill = primary,
			 color = secondary) +
	labs(title = name,
		 x = "Distance from Net",
		 y = "Probability of Scoring") +
	theme_minimal()
	plot
}

make.knn.plot.angle = function(data, primary, secondary, team) {
	name = paste(team, "Predicted Goal Probability from KNN Model", sep = " ")
	plot = ggplot(data) +
	geom_hex(aes(x = angle, y = predict, alpha = ..count..),
			 fill = primary,
			 color = secondary) +
	labs(title = name,
		 x = "Angle of Shot on Net",
		 y = "Probability of Scoring") +
	theme_minimal()
	plot
}

make.nnet.plot.dist = function(data, primary, secondary, team) {
	name = paste(team, "Predicted Goal Probability from NNet Model", sep = " ")
	plot = ggplot(data) +
	geom_hex(aes(x = dist, y = predict, alpha = ..count..),
			 fill = primary,
			 color = secondary) +
	labs(title = name,
		 x = "Distance from Net",
		 y = "Probability of Scoring") +
	theme_minimal()
	plot
}

make.nnet.plot.angle = function(data, primary, secondary, team) {
	name = paste(team, "Predicted Goal Probability from NNet Model", sep = " ")
	plot = ggplot(data) +
	geom_hex(aes(x = angle, y = predict, alpha = ..count..),
			 fill = primary,
			 color = secondary) +
	labs(title = name,
		 x = "Angle of Shot on Net",
		 y = "Probability of Scoring") +
	theme_minimal()
	plot
}
```

Let's take a look at the NHL in it's entirety.

```{r visuals}
plot.nnet.dist = make.nnet.plot.dist(nnet.prediction.data,
									 "orange",
									 "black",
									 "NHL")
plot.nnet.angle = make.nnet.plot.angle(nnet.prediction.data,
									   "orange",
									   "black",
									   "NHL")

plot.knn.dist = make.knn.plot.dist(knn.prediction.data,
								   "orange",
								   "black",
								   "NHL")
plot.knn.angle = make.knn.plot.angle(knn.prediction.data,
									 "orange",
									 "black",
									 "NHL")
```

Here is our neural net model:

```{r vis_nnet_dist}
plot.nnet.dist
```

```{r vis_nnet_angle}
plot.nnet.angle
```

Here is our knn model:

```{r vis_knn_dist}
plot.knn.dist
```

```{r vis_knn_angle}
plot.knn.angle
```

# Analysis {.tabset .tabset-fade}

## Pittsburgh Penguins

Let's first get their data.

```{r pit_data}
pit.nnet = get.team.data(nnet.prediction.data, "PIT")
pit.knn = get.team.data(knn.prediction.data, "PIT")
```

Now, let's see how the Penguins fared in our models.

```{r pit_visuals}
pit.plot.nnet.dist = make.nnet.plot.dist(pit.nnet,
										 "#000000",
										 "#FCB514",
										 "Pittsburgh")
pit.plot.nnet.angle = make.nnet.plot.angle(pit.nnet,
										   "#000000",
										   "#FCB514",
										   "Pittsburgh")

pit.plot.knn.dist = make.knn.plot.dist(pit.knn,
									   "#000000",
									   "#FCB514",
									   "Pittsburgh")
pit.plot.knn.angle = make.knn.plot.angle(pit.knn,
										 "#000000",
										 "#FCB514",
										 "Pittsburgh")
```

Here is the neural network plots:

```{r pit_vis_net_dist}
pit.plot.nnet.dist
```

```{r pit_vis_net_angle}
pit.plot.nnet.angle
```

Here is the K nearest neighbors plots:

```{r pit_vis_knn_dist}
pit.plot.knn.dist
```

```{r pit_vis_knn_angle}
pit.plot.knn.angle
```

## Boston Bruins

Let's first get their data.

```{r bos_data}
bos.nnet = get.team.data(nnet.prediction.data, "BOS")
bos.knn = get.team.data(knn.prediction.data, "BOS")
```

Now, let's see how the Bruins fared in our models.

```{r bos_visuals}
bos.plot.nnet.dist = make.nnet.plot.dist(bos.nnet,
										 "#FFB81C",
										 "#000000",
										 "Boston")
bos.plot.nnet.angle = make.nnet.plot.angle(bos.nnet,
										   "#FFB81C",
										   "#000000",
										   "Boston")

bos.plot.knn.dist = make.knn.plot.dist(bos.knn,
									   "#FFB81C",
									   "#000000",
									   "Boston")
bos.plot.knn.angle = make.knn.plot.angle(bos.knn,
										 "#FFB81C",
										 "#000000",
										 "Boston")
```

Here is the neural network plots:

```{r bos_vis_net_dist}
bos.plot.nnet.dist
```

```{r bos_vis_net_angle}
bos.plot.nnet.angle
```

Here is the K nearest neighbors plots:

```{r bos_vis_knn_dist}
bos.plot.knn.dist
```

```{r bos_vis_knn_angle}
bos.plot.knn.angle
```

## Tampa Bay Lightning

Let's first get their data.

```{r tbl_data}
tbl.nnet = get.team.data(nnet.prediction.data, "T.B")
tbl.knn = get.team.data(knn.prediction.data, "T.B")
```

Now, let's see how the Lightning fared in our models.

```{r tbl_visuals}
tbl.plot.nnet.dist = make.nnet.plot.dist(tbl.nnet,
										 "#002868",
										 "#FFFFFF",
										 "Tampa Bay")
tbl.plot.nnet.angle = make.nnet.plot.angle(tbl.nnet,
										   "#002868",
										   "#FFFFFF",
										   "Tampa Bay")

tbl.plot.knn.dist = make.knn.plot.dist(tbl.knn,
									   "#002868",
									   "#FFFFFF",
									   "Tampa Bay")
tbl.plot.knn.angle = make.knn.plot.angle(tbl.knn,
										 "#002868",
										 "#FFFFFF",
										 "Tampa Bay")
```

Here is the neural network plots:

```{r tbl_vis_net_dist}
tbl.plot.nnet.dist
```

```{r tbl_vis_net_angle}
tbl.plot.nnet.angle
```

Here is the K nearest neighbors plots:

```{r tbl_vis_knn_dist}
tbl.plot.knn.dist
```

```{r tbl_vis_knn_angle}
tbl.plot.knn.angle
```

## San Jose Sharks

Let's first get their data.

```{r sjs_data}
sjs.nnet = get.team.data(nnet.prediction.data, "S.J")
sjs.knn = get.team.data(knn.prediction.data, "S.J")
```

Now, let's see how the Sharks fared in our models.

```{r sjs_visuals}
sjs.plot.nnet.dist = make.nnet.plot.dist(sjs.nnet,
										 "#006D75",
										 "#EA7200",
										 "San Jose")
sjs.plot.nnet.angle = make.nnet.plot.angle(sjs.nnet,
										   "#006D75",
										   "#EA7200",
										   "San Jose")

sjs.plot.knn.dist = make.knn.plot.dist(sjs.knn,
									   "#006D75",
									   "#EA7200",
									   "San Jose")
sjs.plot.knn.angle = make.knn.plot.angle(sjs.knn,
										 "#006D75",
										 "#EA7200",
										 "San Jose")
```

Here is the neural network plots:

```{r sjs_vis_net_dist}
sjs.plot.nnet.dist
```

```{r sjs_vis_net_angle}
sjs.plot.nnet.angle
```

Here is the K nearest neighbors plots:

```{r sjs_vis_knn_dist}
sjs.plot.knn.dist
```

```{r sjs_vis_knn_angle}
sjs.plot.knn.angle
```

## Nashville Predators

Let's first get their data.

```{r nsh_data}
nsh.nnet = get.team.data(nnet.prediction.data, "NSH")
nsh.knn = get.team.data(knn.prediction.data, "NSH")
```

Now, let's see how the Predators fared in our models.

```{r nsh_visuals}
nsh.plot.nnet.dist = make.nnet.plot.dist(nsh.nnet,
										 "#FFB81C",
										 "#041E42",
										 "Nashville")
nsh.plot.nnet.angle = make.nnet.plot.angle(nsh.nnet,
										   "#FFB81C",
										   "#041E42",
										   "Nashville")

nsh.plot.knn.dist = make.knn.plot.dist(nsh.knn,
									   "#FFB81C",
									   "#041E42",
									   "Nashville")
nsh.plot.knn.angle = make.knn.plot.angle(nsh.knn,
										 "#FFB81C",
										 "#041E42",
										 "Nashville")
```

Here is the neural network plots:

```{r nsh_vis_net_dist}
nsh.plot.nnet.dist
```

```{r nsh_vis_net_angle}
nsh.plot.nnet.angle
```

Here is the K nearest neighbors plots:

```{r nsh_vis_knn_dist}
nsh.plot.knn.dist
```

```{r nsh_vis_knn_angle}
nsh.plot.knn.angle
```

## Los Angeles Kings

Let's first get their data.

```{r lak_data}
lak.nnet = get.team.data(nnet.prediction.data, "L.A")
lak.knn = get.team.data(knn.prediction.data, "L.A")
```

Now, let's see how the Kings fared in our models.

```{r lak_visuals}
lak.plot.nnet.dist = make.nnet.plot.dist(lak.nnet,
										 "#111111",
										 "#A2AAAD",
										 "Los Angeles")
lak.plot.nnet.angle = make.nnet.plot.angle(lak.nnet,
										   "#111111",
										   "#A2AAAD",
										   "Los Angeles")

lak.plot.knn.dist = make.knn.plot.dist(lak.knn,
									   "#111111",
									   "#A2AAAD",
									   "Los Angeles")
lak.plot.knn.angle = make.knn.plot.angle(lak.knn,
										 "#111111",
										 "#A2AAAD",
										 "Los Angeles")
```

Here is the neural network plots:

```{r lak_vis_net_dist}
lak.plot.nnet.dist
```

```{r lak_vis_net_angle}
lak.plot.nnet.angle
```

Here is the K nearest neighbors plots:

```{r lak_vis_knn_dist}
lak.plot.knn.dist
```

```{r lak_vis_knn_angle}
lak.plot.knn.angle
```