---
title: "NHL Shot Location Machine Learning Model"
author: "Sasank Vishnubhatla"
date: "January 10, 2019"
html_document:
  runtime: shiny
  theme: simplex
  df_print: paged
  dev: 'svg'
pdf_document:
  df_print: kable
  toc: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

Last Update: `r Sys.time()`

GitHub Repository: [sv4u/goalie-and-skater-heat-maps](https://github.com/sv4u/shot-ml-model)

# Introduction

In the NHL, players have their *sweet-spots*. For skaters, it can be the top of a certain circle, or right in-front of the net. For goalies, it could be where their vision is best and where they have the best angle to cut down a shot. All players have these sweet-spots, but it is difficult to analytically say where they are. By using shot location data, we can determine these locations and create models show where goalies and skaters need improvement and where they succeed.

Before we jump in, let's clean up our R environment and also load in some libraries we will be using.

```{r}
rm(list = ls())

library(purrr)
library(ggplot2)
```

# Data Formatting

To start, we need to read in our data. Our data is formatted nicely in CSV format. We have data from the, 2015-2016 season, 2016-2017 season, 2017-2018 season, and 2018-2019 season (up to 1/30/19). This data was downloaded from [MoneyPuck](http://www.moneypuck.com). Let's first start by loading in all three seasons of data:

```{r}
data.2015 = read.csv("data/2015.csv")
data.2016 = read.csv("data/2016.csv")
data.2017 = read.csv("data/2017.csv")
data.2018 = read.csv("data/2018.csv")
```

> Note: this will take a *relatively* long time to compute as the datasets are large. Each dataset contains all shot data (**including** playoffs).

We'll only look at regular season data. The playoffs in the NHL are a beast of their own.

```{r}
get.regular.season = function(data) {
	subset(data, isPlayoffGame == 0)
}

season.2015 = get.regular.season(data.2015)
season.2016 = get.regular.season(data.2016)
season.2017 = get.regular.season(data.2017)
season.2018 = get.regular.season(data.2018)

```

Now that we have our data, we can remove extraneous columns. Here is a table of what columns we are keeping, and what we are renaming them to:

| Old Column | New Column |
| ---------- | ---------- |
| xCordAdjusted | x |
| yCordAdjusted | y |
| goal | goal |
| shotAngleAdjusted | angle |
| goalieNameForShot | goalie_name |
| shooterName | skater_name |
| game_id | game |

Now, here is the R code to do this subsetting of the original dataset.

```{r}
get.helpful.data = function(data) {
	data.frame(x = data$xCordAdjusted,
		   y = data$yCordAdjusted,
		   goal = data$goal,
		   angle = data$shotAngleAdjusted,
		   goalie_name = data$goalieNameForShot,
		   skater_name = data$shooterName,
		   game = data$game_id)
}

analysis.2015 = get.helpful.data(season.2015)
analysis.2016 = get.helpful.data(season.2016)
analysis.2017 = get.helpful.data(season.2017)
analysis.2018 = get.helpful.data(season.2018)
```

Now, we have all the data we need.

# Function Definitions {.tabset .tabset-fade}

## Generic

## Skaters

Let's first start with a function to get data for a specific skater.

```{r}
get.skater.data = function(data, name) {
	subset(data, skater_name == name)
}
```

## Graphing

# Analysis {.tabset .tabset-fade}
