---
title: "NHL Shot based Goal Scoring Model"
author: "Sasank Vishnubhatla"
date: "February 2, 2019"
output:
  html_document:
    runtime: shiny
    theme: simplex
    df_print: paged
    dev: 'svg'
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

Last Update: `r Sys.Date()`

# Libraries

Before we start, let's load a few libraries.

```{r libraries}
rm(list = ls())

set.seed(100)

library(knitr)
library(ggplot2)
library(caret)
library(doParallel)

registerDoParallel(cores = (detectCores() - 1))
```

We register all but one core so we can have a lot of parallelsism when we start training our models.

# Data Loading

Let's read in our data.

```{r read_csv}
data.2015 = read.csv("data/2015.csv")
data.2016 = read.csv("data/2016.csv")
data.2017 = read.csv("data/2017.csv")
data.2018 = read.csv("data/2018.csv")
```

Now, we will only deal with regular season events. So let's remove the playoffs from our datasets.

```{r remove_playoffs}
get.regular.season = function(data) {
	subset(data, isPlayoffGame == 0)
}

season.2015 = get.regular.season(data.2015)
season.2016 = get.regular.season(data.2016)
season.2017 = get.regular.season(data.2017)
season.2018 = get.regular.season(data.2018)
```

Now let's remove extraneous columns. At the end, we will have the following columns (I've changed their names for ease):

| Old Column Name | New Column Name |
| --------------- | --------------- |
| `xCordAdjusted` | `x` |
| `yCordAdjusted` | `y` |
| `shotAngleAdjusted` | `angle` |
| `shotDistance` | `dist` |
| `goal` | `goal` |

```{r get_helpful}
get.helpful.data = function(data) {
	data.frame(x = data$xCordAdjusted,
		   y = data$yCordAdjusted,
		   angle = data$shotAngleAdjusted,
		   dist = data$shotDistance,
		   goal = data$goal)
}

analysis.2015 = get.helpful.data(season.2015)
analysis.2016 = get.helpful.data(season.2016)
analysis.2017 = get.helpful.data(season.2017)
analysis.2018 = get.helpful.data(season.2018)
```

Sometimes, there is incomplete data. Let's just keep all the complete cases and remove the incomplete ones.

```{r remove_incomplete}
analysis.2015 = analysis.2015[complete.cases(analysis.2015),]
analysis.2016 = analysis.2016[complete.cases(analysis.2016),]
analysis.2017 = analysis.2017[complete.cases(analysis.2017),]
analysis.all = rbind(analysis.2017, rbind(analysis.2016, analysis.2015))
analysis.all = analysis.all[complete.cases(analysis.all),]
analysis.2018 = analysis.2018[complete.cases(analysis.2018),]
```

# Creating the Models

With our data, we can start creating models. We'll be creating the following models:

- Neural Network
- K Nearest Neighbors

```{r train_models}
control = trainControl(method = "repeatedcv", number = 5, repeats = 2)

model.nnet = train(goal ~ . -goal,
                   data = analysis.all,
                   method = "nnet",
                   trControl = control)

model.knn = train(goal ~ . -goal,
				  data = analysis.all,
				  method = "knn",
				  trControl = control)
```

# Extracting Predictions

Our predictions will come from `analysis.2018`. Here's what a little bit of that data looks like:

```{r testing_data}
kable(analysis.2018[1:10,], caption = "Regular Season Shot Data from the 2018-2019 Season")
```

Now, we can use the `predict` function to get our predictions.

```{r get_predictions}
nnet.prediction = predict(model.nnet, newdata = analysis.2018)
knn.prediction = predict(model.knn, newdata = analysis.2018)

nnet.prediction.data = data.frame(analysis.2018)
nnet.prediction.data$predict = nnet.prediction

knn.prediction.data = data.frame(analysis.2018)
knn.prediction.data$predict = knn.prediction
```

# Visualizing the Predictions

With our predictions, let's view how they differ.

```{r visuals}
plot.nnet = ggplot(nnet.prediction.data) +
	geom_hex(aes(x = dist, y = predict, alpha = ..count..),
			 fill = "red",
			 color = "grey") +
	labs(title = "Predicted Goal Probability from Neural Network Model",
		 x = "Distance from Net",
		 y = "Probability of Scoring") +
	theme_minimal()

plot.knn = ggplot(knn.prediction.data) +
	geom_hex(aes(x = dist, y = predict, alpha = ..count..),
			 fill = "red",
			 color = "grey") +
	labs(title = "Predicted Goal Probability from K-Nearest Neighbors Model",
		 x = "Distance from Net",
		 y = "Probability of Scoring") +
	theme_minimal()
```

Here is our neural net model:

```{r vis_nnet}
plot.nnet
```

Here is our knn model:

```{r vis_knn}
plot.knn
```